---
title: "Copepod Markdown"
output: html_document
date: "2022-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes
main food source for larval fishes, delta smelt
4 sampling locations, 4 different habitat types, maybe one habitat type is better
calculated growth rates for copepods
counted microzoo plankton samples, available food sources (others have studied chlorophyll only phytoplankton) This will include all microplankton

two samples per location, discarded 2
ended with 6 samples

# variables
samp_ev = location name
exp 
IC - intial controls

sa - short axis
la - long axis
wi - width

ignore pres_fact, vol_set_ml, n_bugs, time_d

# treatments
final - sample bottles after 24 hours
treatments - sample water plus copepods eat for 24 hours
initial

3 replicates per sample

tot_vol_um3
tot_biomass_pgC

# she did
1) calculate biomass


# goal
what kind of food was available and how does that compare to growth rates?

relationship between growth rates and biomass

reproduction rate:

# what needs to be done
ctbm_grptyp_with_site
organisms

consolidate each organism into small and large, by biomass?

feeding rates - how much of the food is in the sample did the copepods eat

clearance rate - volume of water multiplied by natural log ...

```{r}

library("readxl")
rawcount <- read_excel("100_400_RawCount_R.xlsx")

```


























