---
title: "Copepod Markdown"
output: html_document
date: "2022-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes
main food source for larval fishes, delta smelt
4 sampling locations, 4 different habitat types, maybe one habitat type is better
calculated growth rates for copepods
counted microzoo plankton samples, available food sources (others have studied chlorophyll only phytoplankton) This will include all microplankton

two samples per location, discarded 2
ended with 6 samples

## variables
samp_ev = location name
exp 
IC - intial controls

sa - short axis
la - long axis
wi - width

ignore pres_fact, vol_set_ml, n_bugs, time_d

## treatments
final - sample bottles after 24 hours
treatments - sample water plus copepods eat for 24 hours
initial

3 replicates per sample

tot_vol_um3
tot_biomass_pgC

## already done
1) calculate biomass


## goal
what kind of food was available and how does that compare to growth rates?

relationship between growth rates and biomass

reproduction rate:

## what needs to be done
ctbm_grptyp_with_site
organisms

consolidate each organism into small and large, by biomass?

feeding rates - how much of the food is in the sample did the copepods eat

clearance rate - volume of water multiplied by natural log ...

# Analysis
## Read in Data
```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
headers100 <- read.csv(file="100xHeaders.csv")
headers400 <- read.csv("400xHeaders.csv")
voll_mass <- read.csv("Clean Data/vollbio_all_no0.csv")
RawcountComb <- read.csv("Clean Data/RawcountComb.csv")
headersComb <- read.csv("Clean Data/headersComb.csv")
volbio_all_no0 <- read.csv("Clean Data/vollbio_all_no0.csv")
ct_bmn_fin <- read.csv("Clean Data/ct_bmn_fin.csv")

# For now use this dataframe for everything
ctmb_grptype <- read.csv("Clean Data/ctbmn_grptype.csv")

# or this one
ctbmn_grptyp_mnct <- read_excel("ctbmn_grptyp_mnct.xlsx")

# for mean and other value coutns graph
grpsz_reps <- read.csv(file="Grpsz_reps.csv")
ctsum_rep_mn <- read.csv(file="ctsum_rep_mn_no0.csv")

# combine these with a key
#Rawcount400 <- read_excel("400x_RawCount_R.xlsx")
#Rawcount100 <- read_excel("100x_RawCount_R.xlsx")
#Rawcount100$zoom = 100
#Rawcount400$zoom = 400
#RawcountComb <- rbind(Rawcount100, Rawcount400)

#headers100$zoom = 100
#headers400$zoom = 400
#headersComb <- rbind(headers100, headers400)

# barplot

bp_data <- read.csv("x.csv")

```


# Barplot Data
## how to get the the confidence interval
```{r}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

data_summary(bp_data, varname = totalCPM, groupnames = c(seq, sampEvent))

```

## draw plot gplot
```{r}

df <- bp_data %>% filter(totalCPM != 0)

# create bar plot with reodering of y-axis variable
# use scales parameter to remove empty variable from y-axis
ggplot(df, aes(y=reorder_within(seq,totalCPM, sampEvent), x=totalCPM, fill=seq))+
  geom_col()+
  facet_wrap(~sampEvent, scales= "free", drop =T)+
    scale_y_discrete(labels =  with(df,     as.character(seq)))+
  labs(y="", x="")

```

## base R
Split data into different groups
```{r}
# colors for graph
bp_data$cols = as.numeric(as.factor(bp_data$group))
bp_data <- bp_data[order(bp_data$cols,decreasing = TRUE),]
bp_data$cols_vir = rep(viridis(39), each= 6)

# filter out 0s
df = bp_data %>% filter(totalCPM != 0)

# subset into groups
SJR2 = subset(df, sampEvent == "SJR2")
YBP1 = subset(df, sampEvent == "YBP1")
LSZ2 = subset(df, sampEvent == "LSZ2")
SJR1 = subset(df, sampEvent == "SJR1")
YBP2 = subset(df, sampEvent == "YBP2")
WLD2 = subset(df, sampEvent == "WLD2")

```



```{r}
# order all of the datasets
SJR1 <- SJR1[order(SJR1$totalCPM,decreasing = TRUE),]
SJR2 <- SJR2[order(SJR2$totalCPM,decreasing = TRUE),]
YBP2 <- YBP2[order(YBP2$totalCPM,decreasing = TRUE),]
YBP1 <- YBP1[order(YBP1$totalCPM,decreasing = TRUE),]
LSZ2 <- LSZ2[order(LSZ2$totalCPM,decreasing = TRUE),]
WLD2 <- WLD2[order(WLD2$totalCPM,decreasing = TRUE),]

pdf("C:/Users/dylan/OneDrive/Desktop/Copepods/Copepods/Figures/Samp Event Barplot.pdf",
    height = 10,
    width = 20)
# now plot all of them
par(mfrow = c(2,4))


# SJR1
barplot(SJR1$totalCPM,names.arg = SJR1$seq, col = SJR1$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim = c(0,301), main="SJR1")

# LSZ2
barplot(LSZ2$totalCPM,names.arg = LSZ2$seq, col = LSZ2$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim = c(0,350), main="LSZ2")

# YBP2
barplot(YBP2$totalCPM,names.arg = YBP2$seq, col = YBP2$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim = c(0,250), main="YBP2")

# legend only 1 time
plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n', xlab="", ylab="")
legend("center", legend = unique(bp_data$group), pch=16, col= unique(bp_data$cols_vir), ncol=2)


# SJR2
barplot(SJR2$totalCPM,names.arg = SJR2$seq, col = SJR2$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim = c(0, 120), main= "SJR2")

# WLD2
barplot(WLD2$totalCPM,names.arg = WLD2$seq, col = WLD2$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim = c(0,140), main="WLD2")

# YBP1
barplot(YBP1$totalCPM,names.arg = YBP1$seq, col = YBP1$cols_vir, las =2, cex.lab=1, cex.axis =1, col.axis="black", ylim= c(0, 400), main="YBP1")



#par(mfrow = c(1,1))
dev.off()
```


######################################################################################
# Below Before data was changed

# run with other dataframe
```{r}
ctbmn_grptyp_mnct_grped <- ctbmn_grptyp_mnct %>%
  group_by(grp_typ, exp, samp_ev)%>%
  summarise_at(vars(tot_ct), list(average = mean))

# log values
#grouped_ctbmn_grptyp$log_av <- log(grouped_ctbmn_grptyp$average)

# Split into sampling events
SJR2 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "SJR2")
LSZ2 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "LSZ2")
WLD2 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "WLD2")
YBP1 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "YBP1")
YBP2 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "YBP2")
SJR1 = subset(ctbmn_grptyp_mnct_grped, samp_ev == "SJR1")

ggplot(data= YBP1, aes(x= reorder(grp_typ, -average, FUN = sum), y=average, color=exp))+
  geom_point(stat="identity", position='identity', size = 3)+
  geom_smooth(formula = y~x, method = "lm")+
#  coord_flip(ylim=c(0,75))+
  labs(x="", title = "YBP1 Mean Counts")+
  theme(axis.text.x = element_text(angle = 90), axis.text.y = element_text(size=6))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))+
  scale_y_continuous(name = "Avaerage of Total Counts", breaks = seq(0,1080, 50))

```

# replicates and mean geom point
```{r}
SJR2_reps = subset(grpsz_reps, samp_ev == "SJR2")
SJR2 = subset(ctbmn_grptyp_mnct, samp_ev == "SJR2")
SJR2 <- SJR2[!(SJR2$exp == "IC"), ]
SJR2$grp_sz = SJR2$category

SJR2_reps <- subset(SJR2_reps, grp_sz %in% SJR2$grp_sz)

# make quality pdf file
pdf(file = "Figures/SJR2 means and replicates.pdf",
    width =10,
    height = 8)

ggplot(data= SJR2, aes(x= reorder(grp_sz, mn_ct, 
                                  FUN = sum, decreasing = TRUE), 
                       y=mn_ct, color=exp))+
  geom_point(stat="identity", position='identity', size = 3)+
  labs(x="", title = "SJR2 Means and Replicates Counts")+
  theme_bw()+
  theme(axis.text.x = element_text(), 
       axis.text.y = element_text(size=6))+
  scale_color_manual(values=c('#E69F00','#56B4E9'))+
  #scale_y_continuous(name = "Mean of Total Counts", breaks = seq(0,1080, 50))
  scale_y_log10(name = "Mean Counts with Replicate Counts")+
  geom_point(data=SJR2_reps, aes(x= grp_sz, y = (counts)),
             shape = 1, size=2)+
  coord_flip()
dev.off()
```

SJR1
```{r}
SJR1_reps = subset(grpsz_reps, samp_ev == "SJR1")
SJR1 = subset(ctbmn_grptyp_mnct, samp_ev == "SJR1")
SJR1 <- SJR1[!(SJR1$exp == "IC"), ]
SJR1$grp_sz = SJR1$category

SJR1_reps <- subset(SJR1_reps, grp_sz %in% SJR1$grp_sz)

# make quality pdf file
pdf(file = "Figures/SJR1 means and replicates.pdf",
    width =10,
    height = 8)

ggplot(data= SJR1, aes(x= reorder(grp_sz, mn_ct, 
                                  FUN = sum, decreasing = TRUE), 
                       y=mn_ct, color=exp))+
  geom_point(stat="identity", position='identity', size = 3)+
  labs(x="", title = "SJR1 Means and Replicates Counts")+
  theme_bw()+
  theme(axis.text.x = element_text(), 
       axis.text.y = element_text(size=6))+
  scale_color_manual(values=c('#E69F00','#56B4E9'))+
  #scale_y_continuous(name = "Mean of Total Counts", breaks = seq(0,1080, 50))
  scale_y_log10(name = "Mean Counts with Replicate Counts")+
  geom_point(data=SJR1_reps, aes(x= grp_sz, y = (counts)),
             shape = 1, size=2)+
  coord_flip()
dev.off()
```


# boxplot version of above
```{r}
pdf(file = "Figures/SJR2 boxplots.pdf",
    width =15,
    height = 8)
ggplot(SJR2_reps, aes(x= reorder(grp_sz, counts, FUN = sum, decreasing = T), y = counts, fill=exp))+
  geom_boxplot(position = position_dodge(1))+
  labs(x="", title = "SJR2 Boxplots")+
  coord_flip()
dev.off()
```


## Microplankton ambient analysis
## Using the cbmn group data make the graph with 39 groups total counts on y and separate graphs for each sampling event
#### Split data into sampling events
```{r}
library(tidyverse)
grouped_ctbmn_grptyp <- ctmb_grptype %>%
  group_by(grp_typ, exp, samp_ev)%>%
  summarise_at(vars(tot_ct), list(average = mean))

# make mean and log values
grouped_ctbmn_grptyp$log_av <- log(grouped_ctbmn_grptyp$average)

# Split into sampling events
LSZ2 = subset(grouped_ctbmn_grptyp, samp_ev == "LSZ2")
WLD2 = subset(grouped_ctbmn_grptyp, samp_ev == "WLD2")
YBP1 = subset(grouped_ctbmn_grptyp, samp_ev == "YBP1")
YBP2 = subset(grouped_ctbmn_grptyp, samp_ev == "YBP2")
SJR1 = subset(grouped_ctbmn_grptyp, samp_ev == "SJR1")
SJR2 = subset(grouped_ctbmn_grptyp, samp_ev == "SJR2")
```

####SJR1 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= SJR1, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(stat="identity", position='identity')+
  geom_smooth(formula = y~x, method = "lm")+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))

```

####SJR2 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= SJR2, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(position='stack', stat='identity')+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))+
  theme(legend.position="top")

```

####LSZ2 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= LSZ2, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(position='stack', stat='identity')+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))

```

####WLD2 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= WLD2, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(position='stack', stat='identity')+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))

```

####YBP1 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= YBP1, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(position='stack', stat='identity')+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))

```

####YBP1 stacked barchart with experiments
```{r}
library(ggplot2)
ggplot(data= YBP2, aes(x= reorder(grp_typ, log_av, FUN = sum), y=log_av, color=exp))+
  geom_point(position='stack', stat='identity')+
#  coord_flip(ylim=c(0,75))+
  labs(x="", y="Log Average of Counts")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))

```


### Manipulate data according to idea in overview
*no type column in ctbmn_grptyp*
```{r}
# create long row with all these different values
ctbmn_grptyp$category <- paste(ctbmn_grptyp$grp_typ, ctbmn_grptyp$sa, ctbmn_grptyp$la)

# take the mean value
ctbmn_grptyp$mean_ct <- ctbmn_grptyp$tot_ct / 3

# why is there multiple values group by and summarize
# appears all the values are unique
library(tidyverse)
grouped_ctbmn_grptyp <- ctbmn_grptyp %>%
  group_by(category, exp, samp_ev)%>%
  summarise_at(vars(mean_ct), list(name = mean))

```

#### take a look at what groups we can aggregate
```{r}
# mean is around 50 very wide spread data
boxplot(ctbmn_grptyp$bmn_pgml, ylim=c(0,100))

# categorize data by 4 groups
# https://rcompanion.org/handbook/E_05.html
Percentile_00  = min(ctbmn_grptyp$bmn_pgml)
Percentile_33  = quantile(ctbmn_grptyp$bmn_pgml, 0.33333)
Percentile_67  = quantile(ctbmn_grptyp$bmn_pgml, 0.66667)
Percentile_100 = max(ctbmn_grptyp$bmn_pgml)

RB = rbind(Percentile_00, Percentile_33, Percentile_67, Percentile_100)

dimnames(RB)[[2]] = "Value"

RB
ctbmn_grptyp$size_cat = NA

ctbmn_grptyp$size_cat[ctbmn_grptyp$bmn_pgml >= Percentile_00 & ctbmn_grptyp$bmn_pgml <  Percentile_33]  = "small"

ctbmn_grptyp$size_cat[ctbmn_grptyp$bmn_pgml >= Percentile_33 & ctbmn_grptyp$bmn_pgml <  Percentile_67]  = "medium"

ctbmn_grptyp$size_cat[ctbmn_grptyp$bmn_pgml >= Percentile_67 & ctbmn_grptyp$bmn_pgml <= Percentile_100] = "large"

# could divide this up into 2 categories or maybe graph the types of individuals by size
ggplot(data= ctbmn_grptyp, aes(x= reorder(exp, mean_ct, FUN = sum), y=mean_ct, fill=size_cat))+
  geom_bar(position='stack', stat='identity')+
  coord_flip()+
  labs(x="", y="Mean Count")


```



## USing Volbio data
### Counts by groups
```{r}
library(ggplot2)
#unique(volbio_all_no0$Group)
# about 10 groups already
ggplot(volbio_all_no0, aes(Group, counts, color=Group))+
  geom_boxplot()+
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())

```

###  biomass by groups
```{r}
ggplot(volbio_all_no0, aes(Group, tot_biomass_pgC, color=Group))+
  geom_boxplot()+
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
    coord_cartesian(ylim = c(0, 200000))
# ran this zoomed in which doesnt show top outliers but better representation of true data


```

### Counts by groups percent of each sampling event
```{r}
ggplot(volbio_all_no0, aes(fill=Group, y=counts, x= samp_ev))+
  geom_bar(position='fill', stat='identity')

# also try postion = 'stack' for a stacked barchart
```

### Biomass by groups percent of each sampling event
```{r}
ggplot(volbio_all_no0, aes(fill=Group, y=tot_biomass_pgC, x= samp_ev))+
  geom_bar(position='fill', stat='identity')
```

### Biomass by groups percent of each experimental sample
```{r}
# sort by factoring to get desired ordering
volbio_all_no0$exp <- factor(volbio_all_no0$exp, levels = c("FC", "IC","T24", "site"))

ggplot(volbio_all_no0, aes(fill=Group, y=tot_biomass_pgC, x= exp))+
  geom_bar(position='fill', stat='identity')
```
*Note: maybe we can remove unidentified for this*

#### percentage
```{r}
# this version has no unidentified
volbio_all_id <- subset(volbio_all_no0, Group!="unidentified")

ggplot(volbio_all_id, aes(fill=Group, y=tot_biomass_pgC, x= exp))+
  geom_bar(position='fill', stat='identity')

```

#### stacked
```{r}
# this version has no unidentified
volbio_all_id <- subset(volbio_all_no0, Group!="unidentified")

ggplot(volbio_all_id, aes(fill=Group, y=tot_biomass_pgC, x= exp))+
  geom_bar(position='stack', stat='identity')

```

###  Counts by groups percent of each experimental sample

```{r}
ggplot(volbio_all_no0, aes(fill=Group, y=counts, x= exp))+
  geom_bar(position='fill', stat='identity')
```

#### percentage
```{r}
ggplot(volbio_all_id, aes(fill=Group, y=counts, x= exp))+
  geom_bar(position='fill', stat='identity')

```

#### stacked
```{r}
ggplot(volbio_all_id, aes(fill=Group, y=counts, x= exp))+
  geom_bar(position='stack', stat='identity')

```

### Counts and biomass by groups by sampling event by experimental sample
*Could potentially do this by doing a grid of experimental samples*

### Counts and biomass per volume

### Aggregate organism categories

### Count and biomass per volume aggregated

*This may no be possible:*
## Counts and biomass by groups by sampling event by experimental sample by
replicate.

There is just too much information to display here

# Other graphs of organisms by counts by sampling event separate graphs

### Counts by groups percent of each sampling event
```{r}
# Split into sampling events
LSZ2 = subset(volbio_all_no0, samp_ev == "LSZ2")
WLD2 = subset(volbio_all_no0, samp_ev == "WLD2")
YBP1 = subset(volbio_all_no0, samp_ev == "YBP1")
YBP2 = subset(volbio_all_no0, samp_ev == "YBP2")
SJR1 = subset(volbio_all_no0, samp_ev == "SJR1")
SJR2 = subset(volbio_all_no0, samp_ev == "SJR2")
```

#### LSZ2
```{r}
ggplot(LSZ2, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

#### WLD2
```{r}
ggplot(WLD2, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

#### YBP1
```{r}
ggplot(YBP1, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

#### YBP2
```{r}
ggplot(YBP2, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

#### SJR1
```{r}
ggplot(SJR1, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

#### SJR2
```{r}
ggplot(SJR2, aes(fill=Group, y=counts, x= Group))+
  geom_bar(stat='identity')+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

```

# with the ct_bmn_fin data
```{r}

# Split into sampling events
LSZ2 = subset(ct_bmn_fin, samp_ev == "LSZ2")
WLD2 = subset(ct_bmn_fin, samp_ev == "WLD2")
YBP1 = subset(ct_bmn_fin, samp_ev == "YBP1")
YBP2 = subset(ct_bmn_fin, samp_ev == "YBP2")
SJR1 = subset(ct_bmn_fin, samp_ev == "SJR1")
SJR2 = subset(ct_bmn_fin, samp_ev == "SJR2")

```

## LSZ2 example with log scale
```{r}
p1 <- ggplot(LSZ2, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

## WLD2 example with log scale
```{r}
p2 <- ggplot(WLD2, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

## YBP1 example with log scale
```{r}
p3 <- ggplot(YBP1, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

```

## YBP2 example with log scale
```{r}
p4 <- ggplot(YBP2, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

## SJR1 example with log scale
```{r}
p5 <- ggplot(SJR1, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

## SJR2
```{r}
p6 <- ggplot(SJR2, aes(fill=Group, y=log_mean_ct, x= reorder(Group, -log_mean_ct)))+
  geom_boxplot() + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
```

## all together
```{r}
library(gridExtra)
grid.arrange(p1,p2,p3,p4,p5,p6, nrow=3)
```



# List of the organisms in FC that were not in T24
```{r}
# FC
FC <- RawcountComb[grepl("FC", colnames(RawcountComb))]
FC <- cbind(FC, total = rowSums(FC, na.rm=T))
FC <- cbind(FC, RawcountComb$name)
FC <- data.frame(FC$`RawcountComb$name`, FC$total)
names(FC)[1] <- "org"
names(FC)[2] <- "total"
FC[FC==0]<-NA
FC <- FC[complete.cases(FC),]
FC <- FC$org

# T24
T24 <- RawcountComb[grepl("T24", colnames(RawcountComb))]
T24 <- cbind(T24, total = rowSums(T24, na.rm=T))
T24 <- cbind(T24, RawcountComb$name)
T24 <- data.frame(T24$`RawcountComb$name`, T24$total)
names(T24)[1] <- "org"
names(T24)[2] <- "total"
T24[T24==0]<-NA
T24 <- T24[complete.cases(T24),]
T24 <- T24$org

# Now there are two lists return the list of unmatched
print(FC[!(FC %in% T24)])
# 106 species

```

## barchart of groups by first part of their names
```{r}
library(stringr)
unmatched <- FC[!(FC %in% T24)]
bar_un <- data.frame(unmatched)
bar_un$unknown <- str_count(bar_un$unmatched,"un")
bar_un$diatom <- str_count(bar_un$unmatched,"dia")
bar_un$dinoflaggellate <- str_count(bar_un$unmatched,"dino")
bar_un$cyanobacteria <- str_count(bar_un$unmatched,"cya")
bar_un$ciliate <- str_count(bar_un$unmatched,"cil")
bar_un$chlorophyte <- str_count(bar_un$unmatched,"chl")
bar_un$flagellate <- str_count(bar_un$unmatched,"fla")
bar_un$ochrophyre <- str_count(bar_un$unmatched,"och")
bar_un$tintinnid <- str_count(bar_un$unmatched,"tin")
# not sure about diatom cylindrotheca

# remove column with names and sum
bar_un <- bar_un[,-1]
bar_sums <- colSums(bar_un)
bar_sums <- data.frame(bar_sums)
bar_sums <- cbind(group = rownames(bar_sums), bar_sums)

# make graph
ggplot(data=bar_sums, aes(x=group, y=bar_sums, color=group))+
  geom_bar(stat="identity")+
  theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())

```


# Potential Anaylses:
## PCA
```{r}
library(dplyr)
# Drop columns
spp_mat <- subset(RawcountComb, select = -c(wi, la,sa,shp,name,type,Group, X, X.1, zoom))

# set index and drop row
rownames(spp_mat) = spp_mat$Organism

spp_mat <- subset(spp_mat, select = -c(Organism))

# make it a matrix 
spp_mat <- as.matrix(spp_mat)
# transpose
spp_mat = t(spp_mat)

# make a vector of clean values for locations
location <- gsub(".site", "", rownames(spp_mat))
location <- gsub(".FC.1", "", location)
location <- gsub(".FC.2", "", location)
location <- gsub(".FC.3", "", location)
location <- gsub(".T24.1", "",location)
location <- gsub(".T24.2", "", location)
location <- gsub(".T24.3", "", location)
location <- gsub(".IC.1", "", location)
location <- gsub(".IC.2", "", location)
location <- gsub(".IC.3", "", location)

# bind with matrix
spp_mat <- cbind(spp_mat, location)

# Make log
library(vegan)
# make graph
library(plotly)
library(ggfortify)
library(factoextra)

# Potentially create different ecosystem colors, or different classification
spp_mat[is.na(spp_mat)] <- 0

# remove category
spp_mat_sh_g <- spp_mat[,0:338]

class(spp_mat_sh_g) <- "numeric"

spp_mat_sh_g <- log1p(spp_mat_sh_g)

pca_res <- prcomp(spp_mat_sh_g, scale. = F)

p <- autoplot(pca_res, data=spp_mat, colour= 'location', labels=F, shape=T)
ggplotly(p)


```

# Maps

# FR Groups Plots
```{r}
load("C:/Users/dylan/OneDrive/Desktop/Copepods/Copepods/FR_Groups/FrGrps.Rdata")

library(readxl)
library(ggplot2)
FrGrps <- read_excel("FR_Groups/FrGrps.xlsx")

```


## Things to do
1) Numbers that fall below zero are a different color (done)
2) FrmnCpm indpeendent and second FMnBpm on y axis
3) Negative numbers don't show up, or if they are really close to 0 (done)
4) Shift x axis labels so that they are more to the left end of word going up to line. (Done)
5) Separate into separate plots, going to have to split (done)

Original Code
```{r}
p <- 	ggplot(FrGrps, aes(x=group_size, FRmnCpm))+
   		geom_bar(stat = "identity", aes(fill = FRmnCpm <= 0))+
      scale_colour_manual(name = "FRmnCpm", values = c('TRUE' = "pink", 'FALSE' = "cyan"))+
   		scale_x_discrete ("") +
   		scale_y_continuous(trans = "log10") +
      theme(axis.text.x = element_text(angle = 60, hjust = 0.8, vjust = 0.8, size = 12),
         	strip.text.x = element_text(size = 14))+
   		facet_wrap(~ event, ncol= 2, scales="free") +
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")

p

```

Add a column that is color depending on if value below 0
```{r}
FrGrps$color = "True"
FrGrps$color[FrGrps$FRmnCpm <= 0 ] = "False"

```

Split data
```{r}
# subset and transform into groups
SJR2 = subset(FrGrps, event == "SJR2")
SJR2$tfcpm <- log10(SJR2$FRmnCpm + (-1 * min(SJR2$FRmnCpm)) + 1.1)
SJR2$tfbpm <- log10(SJR2$FRmnBpm + (-1 * min(SJR2$FRmnBpm)) + 1.1)

YBP1 = subset(FrGrps, event == "YBP1")
YBP1$tfcpm <- log10(YBP1$FRmnCpm + (-1 * min(YBP1$FRmnCpm)) + 1.1)
YBP1$tfbpm <- log10(YBP1$FRmnBpm + (-1 * min(YBP1$FRmnBpm)) + 1.1)

LSZ2 = subset(FrGrps, event == "LSZ2")
LSZ2$tfcpm <- log10(LSZ2$FRmnCpm + (-1 * min(LSZ2$FRmnCpm)) + 1.1)
LSZ2$tfbpm <- log10(LSZ2$FRmnBpm + (-1 * min(LSZ2$FRmnBpm)) + 1.1)

SJR1 = subset(FrGrps, event == "SJR1")
SJR1$tfcpm <- log10(SJR1$FRmnCpm + (-1 * min(SJR1$FRmnCpm)) + 1.1)
SJR1$tfbpm <- log10(SJR1$FRmnBpm + (-1 * min(SJR1$FRmnBpm)) + 1.1)

YBP2 = subset(FrGrps, event == "YBP2")
YBP2$tfcpm <- log10(YBP2$FRmnCpm + (-1 * min(YBP2$FRmnCpm)) + 1.1)
YBP2$tfbpm <- log10(YBP2$FRmnBpm + (-1 * min(YBP2$FRmnBpm)) + 1.1)

WLD2 = subset(FrGrps, event == "WLD2")
WLD2$tfcpm <- log10(WLD2$FRmnCpm + (-1 * min(WLD2$FRmnCpm)) + 1.1)
WLD2$tfbpm <- log10(WLD2$FRmnBpm + (-1 * min(WLD2$FRmnBpm)) + 1.1)
```


# FrmnCpm
## SJR2
```{r}
pdf(file = "FR_Groups/Plots/SJR2_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(SJR2, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(SJR2$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("SJR2 FrmnCpm")

dev.off()
```

## SJR1
```{r}
pdf(file = "FR_Groups/Plots/SJR1_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(SJR1, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(SJR1$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("SJR1 FrmnCpm")

dev.off()
```

## WLD2
```{r}
pdf(file = "FR_Groups/Plots/WLD2_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(WLD2, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(WLD2$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("WLD2 FrmnCpm")

dev.off()
```

## YBP1
```{r}
pdf(file = "FR_Groups/Plots/YBP1_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(YBP1, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(YBP1$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("YBP1 FrmnCpm")

dev.off()
```

## YBP2
```{r}
pdf(file = "FR_Groups/Plots/YBP2_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(YBP2, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(YBP2$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("YBP2 FrmnCpm")

dev.off()
```

## LSZ2
```{r}
pdf(file = "FR_Groups/Plots/LSZ2_FrmnCpm.pdf",
    width = 10,
    height = 10)

ggplot(LSZ2, aes(x=group_size,
                 y = tfcpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(LSZ2$tfcpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("LSZ2 FrmnCpm")

dev.off()
```


# FrmBpm
## SJR2
```{r}
pdf(file = "FR_Groups/Plots/SJR2_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(SJR2, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(SJR2$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("SJR2 FrmnBpm")

dev.off()
```

## SJR1
```{r}
pdf(file = "FR_Groups/Plots/SJR1_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(SJR1, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(SJR1$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("SJR1 FrmnBpm")

dev.off()
```

## WLD2
```{r}
pdf(file = "FR_Groups/Plots/WLD2_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(WLD2, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(WLD2$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("WLD2 FrmnBpm")

dev.off()
```

## YBP1
```{r}
pdf(file = "FR_Groups/Plots/YBP1_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(YBP1, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(YBP1$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("YBP1 FrmnBpm")

dev.off()
```

## YBP2
```{r}
pdf(file = "FR_Groups/Plots/YBP2_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(YBP2, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(YBP2$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("YBP2 FrmnBpm")

dev.off()
```

## LSZ2
```{r}
pdf(file = "FR_Groups/Plots/LSZ2_FrmnBpm.pdf",
    width = 10,
    height = 10)

ggplot(LSZ2, aes(x=group_size,
                 y = tfbpm,
                 fill = color))+
   		geom_bar(stat = "identity")+
   		scale_y_continuous(limits = c(0,max(LSZ2$tfbpm)+.5)) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1.15, vjust = 1.1, size = 12),
         	strip.text.x = element_text(size = 14))+
   		xlab("Taxa Groups with Sizes") +
   		ylab("Feeding Rate, Counts per ml")+
      guides(fill = guide_legend(title="More than 0"))+
      ggtitle("LSZ2 FrmnBpm")

dev.off()
```
















